fsOGTT Pre- and Post-ETI

### 2025-04-17 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
EnVision fsOGTT

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

@@

    pancreatic_status, corrector___4, bmi, fev1,

@@

  mutate(ETI = 0) %>%

@@

    cf_genotype, cf_pancreatic, ETI, average_bmi, fev1_percent,

@@

# Merge
df <- full_join(df, gem)

@@

- Make the plot of number of people with OGTTs pre and post
- If people were on Ivacaftor prior to ETI we can still include those visits as pre-ETI
- Adjust models for age
- Paneled figure with results for glucose, c-peptide, and insulin
- Look at acute insulin response (change in 30 minute c peptide)/(change in 30 minute glucose)
  - Can also do across the whole time period and can use insulin in the numerator also
- See Teams folder document for detailed plan

# Table

```{r}
#| label: tbl-planets
#| tbl-cap: Planets
```

# Figure

```{r}
#| label: fig-planets
#| fig-cap: Planets
```

>>>>

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(lmerTest)
library(performance)
library(gtsummary)
library(emmeans)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/EnVision CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/EnVision CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# Get pre-/post-ETI
df$eti_start <- ifelse(df$corrector___4 == 1, df$corrector_start_date, NA)
  group_by(study_id) %>%
  fill(eti_start, .direction = "downup")
df$pre_post <-
  factor(df$Date >= df$eti_start, levels = c(F, T), labels = c("Pre", "Post"))

@@

    pancreatic_status, corrector___4, pre_post, bmi, fev1,

@@

  mutate(ETI = 0, pre_post = "Pre") %>%

@@

    cf_genotype, cf_pancreatic, ETI, pre_post, average_bmi, fev1_percent,

@@

# Merge, keep only pre/post
df <- full_join(df, gem) %>% filter(!is.na(pre_post))
df$pre_post <- factor(df$pre_post, levels = c("Pre", "Post"))

@@

# Plot

```{r}
df %>%
  group_by(study_id) %>%
  count(pre_post) %>%
  ggplot(., aes(y = study_id, x = n)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  facet_wrap(~pre_post)
```

- Adjust models for age
- Paneled figure with results for glucose, c-peptide, and insulin
- Look at acute insulin response (change in 30 minute c peptide)/(change in 30 minute glucose)
  - Can also do across the whole time period and can use insulin in the numerator also
- See Teams folder document for detailed plan

# Glucose

```{r}
# Make a long DF for models
long_df <- df %>% pivot_longer(Glucose_0:Glucose_180,
  names_to = c(".value", "timepoint"),
  names_pattern = "(.*)_(.*)",
  values_to = "Glucose"
)
long_df$timepoint <- factor(long_df$timepoint,
  levels = c("0", "10", "30", "60", "90", "120", "150", "180")
)
mod1 <- lmer(
  Glucose ~ timepoint * pre_post + age_visit + (1 | study_id),
  data = long_df
)
mod1 %>% tbl_regression(.)
means <- data.frame(emmeans(mod1, ~ timepoint + pre_post))
knitr::kable(means)
ggplot(means, aes(
  x = as.numeric(as.character(timepoint)), y = emmean,
  color = pre_post, group = pre_post
)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1) +
  geom_line() +
  theme_bw() +
  xlab("Timepoint") +
  ylab("Glucose (mg/dL)") +
  scale_x_continuous(breaks = c(0, 10, 30, 60, 90, 120, 150, 180))
```
